---
title: "In the zone"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{In the zone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures",
  fig.ext = "png",
  dev = "png")
tryCatch({
  Sys.setlocale("LC_ALL", "English")
})
library(ggplot2)
theme_set(theme_light())
```

> This vignette presents some vital aspects of time zones when adding
'hourglass' layers to your plot

```{r setup, message=FALSE, warning=FALSE}
## load required namespaces for this vignette
library(ggplot2)
library(gghourglass)
library(dplyr)
library(lubridate)
```

Whether it's day or night depends on where you are on the Earth's globe.
This is why you need to pass the longitude and latitude when you
wish to decorate your plot with ribbons indicating night (or day) time.

But which hour of the day is it anyway? Well, that of course depends on
the time zone that you use as a reference. The `?bats` dataset, included
with the `gghourglass` package, expresses the date and time in Coordinated
Universal Time (UTC). This makes sense when your observations are close to
the prime meridian, but not so much when you are in a different time zone
(such as the Central Time Zone in the United States).

Let's make this more clear through an example. We start by plotting the provided observations
of bat call sequences:

```{r plot-utc, fig.width=7, fig.height=3, eval=TRUE}
## get example data
data(bats)

## subset example date to the year 2018
bats_sub <- subset(bats, format(RECDATETIME, "%Y") == "2019")

## retrieve monitoring location
lon <- attr(bats, "monitoring")$longitude[1]
lat <- attr(bats, "monitoring")$latitude[1]

## plot the data
p <-
  ggplot(bats_sub, aes(x = RECDATETIME)) +
  
    ## annotate sunset until sunrise
    annotate_daylight(lon, lat, c("sunset", "sunrise")) +
  
    ## annotate dusk until dawn
    annotate_daylight(lon, lat, c("dusk", "dawn")) +
  
    ## add hourglass geometry to plot
    geom_hourglass() +
  
    ## add informative labels
    labs(x = "Date", y = "Time of day")
p
```

Note that the `geom_hourglass()` layer automatically uses the time zone as
specified by the data (UTC). If you wish to display it in a different time
zone, such as for instance Central European Time (CET), you simply mutate
the time zone of the observations with `lubridate::with_tz()`:

```{r plot-cet, fig.width=7, fig.height=3, eval=TRUE, warning=FALSE}
p %+%
  mutate(bats_sub, RECDATETIME = with_tz(RECDATETIME, "CET"))
```

Hold on, why is the ribbon indicating night time jagged? This is because
CET has summer daylight-saving time which causes the night-time to shift
one hour.
